The Spatial and Temporal Domains of Modern Ecology

### 2017-12-22 paper/main.tex
Main and SI submissions, pacakge incremented to 0.4.0

\noindent \textbf{\emph{Distributions within the four primary dimensions}}\\
In terms of resolution, spatial replicates for the majority (67\%) of observations (across all methods) were $<$1 m$^2$, a further 24\% were 1 m$^2$ up to 1 ha, and the remaining 9\% were $\geq$1 ha (Fig. 1A). These distributions primarily reflect those of field observations, the dominant observational methodology. Examining the distributions for each observational method (Fig. S1 in SI) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $<$0.1 m$^2$) than field observations (47\% $<$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $>$100 m$^2$).

The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 12\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig. 1B).  As with resolution, the extent covered by automated sensing methods tended to be smaller (52\% $<$100 ha) than those of field observations (31\% $<$100 ha), while all but 4\% of remote observations covered areas $\geq$10,000 ha (as did the small number of paleo-reconstructions).

@@

In the temporal dimensions, 37\% of observations were not repeated (Fig. 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals. With respect to temporally replicated observation (Fig. S1 in SI), automated sensing techniques had the finest intervals (61\% $\leq$1 day; 100\% $\leq$1 year), followed by remote observation (37\% $\leq$1 day; 78\% $\leq$1 year), field observations (17\% $\leq$1 day; 86\% $\leq$1 year), and paleo-reconstructions (21\% $\leq$1 decade).

@@

\noindent \textbf{\emph{Spatial and temporal domains}}\\
Juxtaposing these observational dimensions provides further insight into the spatial and temporal domains of observations (Fig. 2). Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the upper left by monthly to yearly observations at 100 m$^2$ resolution and on the lower right by near-daily to monthly observations with 1-10 ha resolution is also evident. The four observational methods occupied substantially different portions of the domain space, as indicated by the locations of their median values (and illustrated further in Fig. S2 in the SI): the median domain of field observations was between 0.1-1 m$^2$ of resolution with a monthly interval, whereas remote observations had coarser median resolutions (1000 m$^2$) but finer median intervals ($\sim$1 day). Paleo-reconstructions and automated sensing techniques were both finely resolved (medians from 10 cm$^2$ to 0.01 m$^2$), but automated approaches had hourly-daily median intervals compared to multi-decadal for paleo-reconstructions.

%Comparing the spatial and temporal resolutions of  interval to spatial resolution of ecological observations reveals three primary regions of observational concentration (Fig. 2A), the greatest being unreplicated samples with resolutions of 100 cm$^2$ to 1 m$^2$, followed by observations of daily to monthly interval with slightly finer spatial resolutions (10 cm$^2$ to 1 m$^2$), and finally a smaller concentration of monthly to yearly measurements collected at spatial scales of 0.1-10 ha.

@@

Contrasting the two spatial dimensions against one another (for all observations) shows a primary concentration of observations with 10 cm$^2$ to 100 m$^2$ resolution with extents ranging from just over 1,000 to nearly 1,000,000 ha (Fig. 2C). The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution, 1000 m$^2$ to $<$10 ha. These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Fig. S2 in SI) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.  Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a result that likely reflects the very low sample size for this observational type.

Two primary domains of observational concentration are revealed by juxtaposing duration and extent (across all observations). The first consists of observations spanning one month to one decade in time and 10-1,000 ha in space, while the second is defined by observations of one year to several decades that cover 10,000 to 1,000,000 ha (Fig. 2D).  Three other notable, but lesser areas of concentration are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($<$1 day) of either 1-10 ha or 10,000-1,000,000 ha.  The median observation (1 year duration, 100 ha extent) from automated sensing lies near the center of the second major concentration, while the median extents of field (1-10,000 ha) and remote (1-10 million ha) observations bound the primary concentration at its upper and lower extents, while the median duration of both observational types falls between 1 month to 1 year.

@@

The relatively small sample size of our study may also bias these findings. Although we used a randomized title draw to obtain a representative sample, we reviewed just 0.8\% of papers published during the 10 year period. Our sample may therefore under- or over-represent observational coverage in certain scale domains, particularly for analyses broken down by observational method. The case where this is of most concern is for paleo-reconstructions, the small sample size of which resulted in a likely over-estimate of the typical extent of such observations (e.g. Fig. 2B, C; however, the interval and duration findings are probably more representative).

>>>>

\noindent \textbf{\emph{Distributions within the four primary dimensions}}\\
In terms of resolution, spatial replicates for the majority (67\%) of observations (across all methods) were $<$1 m$^2$, a further 24\% were 1 m$^2$ up to 1 ha, and the remaining 9\% were $\geq$1 ha (Fig. 1A). These distributions primarily reflect those of field observations, the dominant observational methodology. Examining the distributions for each observational method (Fig. S1 in SI) shows that automated sensing and paleo-reconstruction had resolutions that were generally finer (85\% or more $<$0.1 m$^2$) than field observations (47\% $<$0.1 m$^2$), while the majority of remote observations were much coarser (70\% $\geq$100 m$^2$).

The extent of 19\% of observations was $<$10 ha, 23\% covered 10-1,000 ha, 11\% 1,000-10,000 ha, 19\% 10,000-100,000 ha, 12\% 100,000-1,000,000 ha, and 15\% $>$1,000,000 ha (Fig. 1B).  As with resolution, the extent covered by automated sensing methods tended to be smaller (52\% $<$100 ha) than those of field observations (31\% $<$100 ha), while all but 4\% of remote observations covered areas $\geq$10,000 ha (as did the small number of paleo-reconstructions).

@@

In the temporal dimensions, 37\% of observations were not repeated (Fig. 1C), 17\% were repeated at short intervals (sub-second to daily), 20\% at daily to monthly intervals, 18\% at monthly to yearly intervals, 6\% at yearly to decadal intervals, and 2\% at decadal or greater intervals. With respect to temporally replicated observations (Fig. S1 in SI), automated sensing techniques had the finest intervals (61\% $\leq$1 day; 100\% $\leq$1 year), followed by remote observation (37\% $\leq$1 day; 78\% $\leq$1 year), field observations (17\% $\leq$1 day; 86\% $\leq$1 year), and paleo-reconstructions (21\% $\leq$1 decade).

@@

\noindent \textbf{\emph{Spatial and temporal domains}}\\
Juxtaposing these observational dimensions provides further insight into the spatial and temporal domains of observations (Fig. 2). Contrasting resolution with interval reveals that the majority of temporally replicated observations (unrepeated observations were excluded because they lack interval values) had resolutions of 10 cm$^2$-1 m$^2$ and were revisited at daily to yearly intervals (Fig. 2A). A less dense, oblong concentration of observations bounded on the upper left by monthly to yearly observations at 100 m$^2$ resolution and on the lower right by near-daily to monthly observations with 1-10 ha resolution is also evident. The four observational methods occupied substantially different portions of the domain space, as indicated by the locations of their median values (and illustrated further in Fig. S2 in the SI): the median domain of field observations was between 0.1-1 m$^2$ of resolution with a monthly interval, whereas remote observations had a coarser median resolution (1000 m$^2$) but finer median interval ($\sim$1 day). Paleo-reconstructions and automated sensing techniques were both finely resolved (median between10 cm$^2$-0.01 m$^2$), but automated approaches had an hourly-daily median interval compared to multi-decadal for paleo-reconstructions.

@@

Contrasting the two spatial dimensions against one another (for all observations) shows a primary concentration of observations of 10 cm$^2$ to nearly 100 m$^2$ resolution with extents ranging from $\sim$1,000 to nearly 1,000,000 ha (Fig. 2C). The second-most prominent concentration consists of higher resolution (1 cm$^2$-1 m$^2$), smaller extent (10-1,000 ha) observations, beneath which lies a third and fainter concentration of 1-1,000 cm$^2$ resolution and 1000 m$^2$ to $<$10 ha extent. These three concentrations suggest a tendency for observational extent to increase with resolution, which is further evident in the median domain values (and kernel densities; Fig. S2 in SI) of automated (0.01 m$^2$ resolution, 100 ha extent), field (0.1-1 m$^2$ resolution, 1,000-10,000 ha extent), and remote (1,000 m$^2$ resolution, 1-10 million ha extent) observations.  Paleo-reconstructions were an outlier from this relationship, having very fine median resolution (0.01 m$^2$) but large extent (1 million ha), a result that likely reflects the very low sample size for this observational type.

Two primary domains of observational concentration are revealed by juxtaposing duration and extent (across all observations). The first consists of observations spanning just over one month to one decade in time and 10-1,000 ha in space, while the second is defined by observations of one year to several decades that cover 10,000 to 1,000,000 ha (Fig. 2D).  Three other notable, but lesser areas of concentration are also evident, including small area observations (0.1-1 ha) covering one month to decade, and short duration, temporally unreplicated observations ($<$1 day) of either 1-10 ha or 10,000-1,000,000 ha.  The median observation from automated sensing (1 year duration, 100 ha extent) lies near the center of the first major concentration, while the median extents of field (1-10,000 ha) and remote (1-10 million ha) observations bound the second major concentration at its upper and lower extents, with the median duration of both observational types falling between 1 month to 1 year.

@@

The relatively small sample size of our study may also bias these findings. Although we used a randomized title draw to obtain a representative sample, we reviewed just 0.8\% of papers published during the 10 year period. Our sample may therefore under- or over-represent observational coverage in certain scale domains, particularly for analyses broken down by observational method. The case where this is of most concern is for paleo-reconstructions, the small sample size of which resulted in a likely over-estimate of the typical extent of such observations (e.g. Fig. 2B, C; however, the interval and duration values are probably more representative).
